---
title: "hw004_ntjjmak"
author: "Nicole Mak"
date: "07/10/2018"
output: github_document
---

### There are two tasks for this assignment.

1. Data reshaping.
2. Perform a join.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


To start, let's load the packages to be used for data exploration.

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(knitr))
```

## Exercise #1: Reshaping data

The data to be used for the exercise is "gapminder".

Here is the data set in its original form:

```{r}
gapminder
```



To practise reshaping the data, I have selected prompt #2. I have not yet tried `knitr::kable()`, so this is a good chance. Here are the exercise instructions: 

*Make a tibble with one row per year and columns for life expectancy for two or more countries. Use `knitr::kable()` to make this table look pretty in your rendered homework. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.*



**First, let's use `dplyr` to get the information we want to focus on. We will select for the variables (i.e. columns) we want, make sure it is grouped by country, and filter only for the countries desired. This will be stored under "reshape".**
```{r}
reshape <- gapminder %>% 
  select(year, country, lifeExp)%>% 
  group_by(country)%>% 
  filter(country == "Canada" | country == "France" | country == "Japan")
```

**Now we can reformat this so that there is only one row per year and present this data in a table by using  `knitr::kable()`. We will round to the nearest decimal point.**


Activity #2
Create your own cheatsheet patterned after Jenny’s but focused on something you care about more than comics! Inspirational examples: 

```{r}
spread(reshape, key = country, value = lifeExp) %>% 
  knitr:: kable(format = "markdown", justify = "centre", digits = 1, caption = "Life expectancy 1952-2007 for selected countries: Canada, France, and Japan.")


```

**This same data may also be presented nicely with a scatter plot.**

```{r}
reshape %>% 
  ggplot(aes(year, lifeExp))+
  geom_point(aes(colour = country))+
  theme_classic()+
  ggtitle("Life Expectancy from 1952-2007")+
  ylab("Life Expectancy (years)")+
  xlab("Year")
```


**The next exercise will be to explore `join`. Here are the instructions for the exercise: Create your own cheatsheet patterned after Jenny’s but focused on something you care about more than comics! **
