---
title: "hw004_ntjjmak"
author: "Nicole Mak"
date: "07/10/2018"
output: github_document
---

### There are two tasks for this assignment.

1. Data reshaping.
2. Perform a join.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


To start, let's load the packages to be used for data exploration.

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyr))
```

## Exercise #1: Reshaping data

The data to be used for the exercise is "gapminder".

Here is the data set in its original form:

```{r}
gapminder
```



To practise reshaping the data, I have selected prompt #2. I have not yet tried `knitr::kable()`, so this is a good chance. Here are the exercise instructions: 

*Make a tibble with one row per year and columns for life expectancy for two or more countries. Use `knitr::kable()` to make this table look pretty in your rendered homework. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.*