---
title: "hw004_ntjjmak"
author: "Nicole Mak"
date: "07/10/2018"
output: github_document
---

### There are two tasks for this assignment.

1. Data reshaping.
2. Perform a join.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


To start, let's load the packages to be used for data exploration.

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
```

## Exercise #1: Reshaping data

The data to be used for the exercise is "gapminder".

Here is the data set in its original form:

```{r}
gapminder
```



To practise reshaping the data, I have selected prompt #2. I have not yet tried `knitr::kable()`, so this is a good chance. Here are the exercise instructions: 

*Make a tibble with one row per year and columns for life expectancy for two or more countries. Use `knitr::kable()` to make this table look pretty in your rendered homework. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.*



**First, let's use `dplyr` to get the information we want to focus on. We will select for the variables (i.e. columns) we want, make sure it is grouped by country, and filter only for the countries desired. This will be stored under "reshape".**
```{r}
reshape <- gapminder %>% 
  select(year, country, lifeExp)%>% 
  group_by(country)%>% 
  filter(country == "Canada" | country == "France" | country == "Japan")
```

**Now we can reformat this so that there is only one row per year and present this data in a table by using  `knitr::kable()`. We will round to the nearest decimal point.**


```{r}
spread(reshape, key = country, value = lifeExp) %>% 
  knitr:: kable(format = "markdown", justify = "centre", digits = 1, caption = "Life expectancy 1952-2007 for selected countries: Canada, France, and Japan.")


```

**This same data may also be presented nicely with a scatter plot.**

```{r}
reshape %>% 
  ggplot(aes(year, lifeExp))+
  geom_point(aes(colour = country))+
  theme_classic()+
  ggtitle("Life Expectancy from 1952-2007")+
  ylab("Life Expectancy (years)")+
  xlab("Year")
```


### The next exercise will be to explore `join`. Here are the instructions for the exercise: Create your own cheatsheet patterned after Jenny’s but focused on something you care about more than comics!


**First, create data frames. The theme for this is movie stars. Start by making vectors for each variable. This code chunk shows one way to make a data frame:**
```{r}
Name <- c('Harrison Ford', 'Gal Gadot', 'Jackie Chan', 'Emma Thompson', 'Leonardo DiCaprio', 'Emily Blunt', 'Natalie Portman')
Sex <- c('M', 'F', 'M', 'F', 'M', 'F', 'F')
Notable_movie <- c('Indiana Jones', 'Wonder Woman', 'Rush Hour', 'Sense and Sensibility', 'Titanic', 'Sicario', 'Star Wars')

Actor.data <- data.frame(Name, Sex, Notable_movie)
```


**This code chunk shows an alternative way to code a data frame.**

```{r}
Movie.data <- data_frame(
  Notable_movie =  c('Indiana Jones', 'Wonder Woman', 'Rush Hour', 'Sense and Sensibility', 'Titanic', 'Star Wars', 'The Red Violin', 'Avengers'),
  Genre_of_movie = c('Action', 'Superhero', 'Action', 'Romance', 'Romance', 'SciFi', 'Drama', 'Superhero'))
```

**Next, code the data frames and format them as a tibble. We will store them as table "x" and table "y".**

```{r}
x <- as.tibble(Actor.data)
y <- as.tibble(Movie.data)
```

**We can even display the  as nicer tables using  `knitr`**

```{r}
as.tibble(x) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
as.tibble(y) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
```

**Let's try different forms of joining to see how our data is re-formatted.**

**First up, `inner_join`. Inner join takes all matching rows from "x" and adds the additional columns from "y" that correspond to them. Notice that only rows from "x" that had corresponding data from "y" are included. Therefore, Emily Blunt no longer appears in the joined data.**


```{r}
inner_join(x, y) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
```


**Next we have `semi_join`. Again, only rows from "x" with corresponding data in "y" will appear. Furthermore, `semi_join` does not display any columns from "y". It only shows you the data for the common rows between the two merged data frames.**
```{r}
semi_join(Actor.data, Movie.data) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
```


**Let's see what happens when we do `left_join` and `right_join` and compare the two.**

```{r}
left_join(x, y) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
right_join(x, y) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
```

**What do you notice?**

`Left_join` will include all rows of data from "x" and add new columns from "y". Any missing matches in the new column pulled from "y" will be left blank as an *NA*. Such is the case for the movie "Sicario" which does not appear in the "y" data. Furthermore, "The Red Violin" and "Avengers" are not rows in "x" and are not even displayed.

`Right_join` is essentially the opposite of `left_join`. It includes all data from the "y" rows and adds on the columns from "x". Again, any "x" data that does not correspond with "y" is left blank and replaced with an *NA*. Here, we see the "Red Violin" and "Avengers" make their re-appearance while "Emily Blunt" disappears once again because she does not have a corresponding row in "y".


**Let's try `anti_join` which will filter out all rows in "x" that had shared data with "y". We are left with the original "x" columns and only rows that did not have a "y" match.**
```{r}
anti_join(x, y) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
```

**Last, but not least, we have `full_join`. True to its name, this requests that all rows of data and columns from both sets are represented. Missing data is displayed as *NA*.**

```{r}
full_join(x, y) %>% 
  knitr:: kable(format = "markdown", justify = "centre")
```


### All of these "join" functions were from the `dplyr` universe. What about the "merge" function from base R?

**Let's try it.**

```{r}
merge(x, y)
merge(y, x)
```


**What we notice is that the merge function returns the same information as inner_join. However, when requesting the coding for the joining of tables, the sequence of columns on display is not affected by the written code. Check out the code chunk below and notice that `inner_join(x, y)` does not return the same results as `inner_join(y, x)`! That is, the column order is affected by how you code it.**

```{r}
inner_join(x, y)
inner_join(y, x)
```



### Thanks for having a look at this assignment!